---
title: "BSA - Hustle"
author: "Grant Boydell"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r}
hustle <- read.csv("pergamehustle2024-25.csv")
```

## Per Game Leaders

### Deflections

```{r}
hustle_by_def <- hustle[order(hustle$Deflections, decreasing = TRUE), ]
par(mar = c(7, 5, 4, 4))
def_dp <- barplot(height = hustle_by_def$Deflections[1:10], 
        names.arg = hustle_by_def$Player[1:10], cex.names = 0.6, las = 2, 
        ylab = "Deflections", border = "skyblue", 
        col = adjustcolor("skyblue", alpha.f = 0.7))
text(def_dp, 0, hustle_by_def$Deflections[1:10], cex = 0.7, pos = 3)
```

### Loose Balls Recovered

```{r}
hustle_by_lb <- hustle[order(hustle$Loose.Balls.Recovered, decreasing = TRUE),]
lb_bp <- barplot(height = hustle_by_lb$Loose.Balls.Recovered[1:10], 
        names.arg = hustle_by_lb$Player[1:10], cex.names = 0.4, las = 2, 
        ylab = "Loose Balls Recovered", border = "skyblue", 
        col = adjustcolor("skyblue", alpha.f = 0.7))
text(lb_bp, 0, hustle_by_lb$Loose.Balls.Recovered[1:10], cex = 0.7, pos = 3)
```

### Screen Assists 

```{r}
hustle_by_sa <- hustle[order(hustle$Screen.Assists, decreasing = TRUE), ]
par(mar = c(7, 5, 4, 4))
sa_bp <- barplot(height = hustle_by_sa$Screen.Assists[1:10], 
        names.arg = hustle_by_sa$Player[1:10], cex.names = 0.6, las = 2, 
        ylab = "Screen Assists", border = "skyblue", 
        col = adjustcolor("skyblue", alpha.f = 0.7))
text(sa_bp, 0, hustle_by_sa$Screen.Assists[1:10], cex = 0.7, pos = 3)
```

### Contested Shots

```{r}
hustle_by_cs <- hustle[order(hustle$Contested.Shots, decreasing = TRUE), ]
par(mar = c(7, 5, 4, 4))
cs_bp <- barplot(height = hustle_by_cs$Contested.Shots[1:10], 
        names.arg = hustle_by_cs$Player[1:10], cex.names = 0.6, las = 2, 
        ylab = "Contested Shots", border = "skyblue", 
        col = adjustcolor("skyblue", alpha.f = 0.7))
text(cs_bp, 0, hustle_by_cs$Contested.Shots[1:10], cex = 0.7, pos = 3)
```

### Charges Drawn

```{r}
hustle_by_cd <- hustle[order(hustle$Charges.Drawn, decreasing = TRUE), ]
par(mar = c(7, 5, 4, 4))
cd_bp <- barplot(height = hustle_by_cd$Charges.Drawn[1:10], 
        names.arg = hustle_by_cd$Player[1:10], cex.names = 0.6, las = 2, 
        ylab = "Charges Drawn", border = "skyblue", 
        col = adjustcolor("skyblue", alpha.f = 0.7))
text(cd_bp, 0, hustle_by_cd$Charges.Drawn[1:10], cex = 0.7, pos = 3)
```

### Box Outs

```{r}
hustle_by_bo <- hustle[order(hustle$Box.Outs, decreasing = TRUE), ]
par(mar = c(7, 5, 4, 4))
bo_bp <- barplot(height = hustle_by_bo$Box.Outs[1:10], 
        names.arg = hustle_by_bo$Player[1:10], cex.names = 0.6, las = 2, 
        ylab = "Box Outs", border = "skyblue", 
        col = adjustcolor("skyblue", alpha.f = 0.7))
text(cd_bp, 0, hustle_by_bo$Box.Outs[1:10], cex = 0.7, pos = 3)
```

## Hustle Metric

```{r}
library(gt)
library(dplyr)

hustle_metric <- function(deflections, loose, screens, contests, charges, boxouts){
  def <- deflections/max(deflections)
  lbr <- loose/max(loose)
  sa <- screens/max(screens)
  cs <- contests/max(contests)
  cd <- charges/max(charges)
  bo <- boxouts/max(boxouts)
  metric <- (lbr*0.3) + (def*0.3) + (cd*0.25) + (bo*0.05) + (cs*0.05) + (sa*0.05)
  metric * 100
}

player_metric <- hustle_metric(hustle$Deflections,
                               hustle$Loose.Balls.Recovered,
                               hustle$Screen.Assists,
                               hustle$Contested.Shots, hustle$Charges.Drawn,
                               hustle$Box.Outs)

hustle$Hustle.Rating <- player_metric

hustle_by_rating <- hustle[order(hustle$Hustle.Rating, decreasing = TRUE), ]
colnames(hustle_by_rating)[6:11] <- c("Screen Assists", "Deflections", "Loose Balls Recovered", "Charges Drawn", "Contested Shots", "Box Outs")
top_hustlers <- data.frame(hustle_by_rating$Player[1:10], round(hustle_by_rating$Hustle.Rating[1:10], digits = 2))
colnames(top_hustlers) <- c("Player", "Hustle Rating")
gt(top_hustlers) %>%
  tab_header(title = md("**Top Hustlers**"), subtitle = md("2024-2025"))
```

## Hustle Player of the Year? 

### Add Percentiles
```{r}
hustle$def_percentiles <- ecdf(hustle$Deflections)(hustle$Deflections) * 100
hustle$loose_percentiles <-
  ecdf(hustle$Loose.Balls.Recovered)(hustle$Loose.Balls.Recovered) * 100
hustle$charges_percentiles <- 
  ecdf(hustle$Charges.Drawn)(hustle$Charges.Drawn) * 100
hustle$contests_percentiles <-
  ecdf(hustle$Contested.Shots)(hustle$Contested.Shots) * 100
hustle$screen_percentiles <-
  ecdf(hustle$Screen.Assists)(hustle$Screen.Assists) * 100
hustle$box_percentiles <- ecdf(hustle$Box.Outs)(hustle$Box.Outs) * 100
```



```{r}
indices <- c(which(hustle$Player == "Toumani Camara"), 
             which(hustle$Player == "Dyson Daniels"), 
             which(hustle$Player == "Paul George"), 
             which(hustle$Player == "Victor Wembanyama"), 
             which(hustle$Player == "Draymond Green"), 
             which(hustle$Player == "Alperen Sengun"), 
             which(hustle$Player == "Josh Hart"), 
             which(hustle$Player == "Alex Caruso"), 
             which(hustle$Player == "Cason Wallace"), 
             which(hustle$Player == "Jalen Suggs"))

percentile_data <- cbind(hustle$def_percentiles[indices], hustle$loose_percentiles[indices], hustle$screen_percentiles[indices], hustle$charges_percentiles[indices], hustle$contests_percentiles[indices], hustle$box_percentiles[indices])


library(fmsb)

radar_data <- as.data.frame(rbind(rep(100, 6), rep(20, 6), percentile_data))
colnames(radar_data) <- c("Deflections", "Loose Balls", "Screen Assists", "Charges", "Contests", "Box Outs")

radar_data2 <- as.data.frame(rbind(c(max(hustle$Screen.Assists), max(hustle$Deflections), max(hustle$Loose.Balls.Recovered), max(hustle$Charges.Drawn), max(hustle$Contested.Shots), max(hustle$Box.Outs)), c(min(hustle$Screen.Assists), min(hustle$Deflections), min(hustle$Loose.Balls.Recovered), min(hustle$Charges.Drawn), min(hustle$Contested.Shots), min(hustle$Box.Outs)), hustle_by_rating[1:10, 6:11]))


radarchart(radar_data[1:3, ], pcol = "red", pfcol = adjustcolor("red", alpha.f = 0.5), plwd = 2, cglcol = "gray", cglty = 1, vlcex = 0.5, title = "Toumani Camara")
radarchart(radar_data[c(1, 2, 4), ], pcol = "darkred", pfcol = adjustcolor("darkred", alpha.f = 0.5), plwd = 2, cglcol = "gray", cglty = 1, vlcex = 0.5, title = "Dyson Daniels")
radarchart(radar_data[c(1, 2, 5), ], pcol = "darkblue", pfcol = adjustcolor("darkblue", alpha.f = 0.5), plwd = 2, cglcol = "gray", cglty = 1, vlcex = 0.5, title = "Paul George")
radarchart(radar_data[c(1, 2, 6), ], pcol = "black", pfcol = adjustcolor("black", alpha.f = 0.5), plwd = 2, cglcol = "gray", cglty = 1, vlcex = 0.5, title = "Victor Wembanyama")
radarchart(radar_data[c(1, 2, 7), ], pcol = "gold", pfcol = adjustcolor("gold", alpha.f = 0.5), plwd = 2, cglcol = "gray", cglty = 1, vlcex = 0.5, title = "Draymond Green")
radarchart(radar_data[c(1, 2, 8), ], pcol = "red", pfcol = adjustcolor("red", alpha.f = 0.5), plwd = 2, cglcol = "gray", cglty = 1, vlcex = 0.5, title = "Alperen Sengun")
radarchart(radar_data[c(1, 2, 9), ], pcol = "blue", pfcol = adjustcolor("blue", alpha.f = 0.5), plwd = 2, cglcol = "gray", cglty = 1, vlcex = 0.5, title = "Josh Hart")
radarchart(radar_data[c(1, 2, 10), ], pcol = "darkorange", pfcol = adjustcolor("darkorange", alpha.f = 0.5), plwd = 2, cglcol = "gray", cglty = 1, vlcex = 0.5, title = "Alex Caruso")
radarchart(radar_data[c(1, 2, 11), ], pcol = "darkorange", pfcol = adjustcolor("darkorange", alpha.f = 0.5), plwd = 2, cglcol = "gray", cglty = 1, vlcex = 0.5, title = "Cason Wallace")
radarchart(radar_data[c(1, 2, 12), ], pcol = "navy", pfcol = adjustcolor("navy", alpha.f = 0.5), plwd = 2, cglcol = "gray", cglty = 1, vlcex = 0.5, title = "Jalen Suggs")

hustle[which(hustle$Player == "Devin Booker"), ]
```
## Actual stats instead of percentiles

```{r}
radarchart(radar_data2[1:3, ], pcol = "red", pfcol = adjustcolor("red", alpha.f = 0.5), plwd = 2, cglcol = "gray", cglty = 1, vlcex = 0.5, title = "Toumani Camara")
radarchart(radar_data2[c(1, 2, 4), ], pcol = "darkred", pfcol = adjustcolor("darkred", alpha.f = 0.5), plwd = 2, cglcol = "gray", cglty = 1, vlcex = 0.5, title = "Dyson Daniels")
radarchart(radar_data2[c(1, 2, 5), ], pcol = "darkblue", pfcol = adjustcolor("darkblue", alpha.f = 0.5), plwd = 2, cglcol = "gray", cglty = 1, vlcex = 0.5, title = "Paul George")
radarchart(radar_data2[c(1, 2, 6), ], pcol = "black", pfcol = adjustcolor("black", alpha.f = 0.5), plwd = 2, cglcol = "gray", cglty = 1, vlcex = 0.5, title = "Victor Wembanyama")
radarchart(radar_data2[c(1, 2, 7), ], pcol = "gold", pfcol = adjustcolor("gold", alpha.f = 0.5), plwd = 2, cglcol = "gray", cglty = 1, vlcex = 0.5, title = "Draymond Green")
radarchart(radar_data2[c(1, 2, 8), ], pcol = "red", pfcol = adjustcolor("red", alpha.f = 0.5), plwd = 2, cglcol = "gray", cglty = 1, vlcex = 0.5, title = "Alperen Sengun")
radarchart(radar_data2[c(1, 2, 9), ], pcol = "blue", pfcol = adjustcolor("blue", alpha.f = 0.5), plwd = 2, cglcol = "gray", cglty = 1, vlcex = 0.5, title = "Josh Hart")
radarchart(radar_data2[c(1, 2, 10), ], pcol = "darkorange", pfcol = adjustcolor("darkorange", alpha.f = 0.5), plwd = 2, cglcol = "gray", cglty = 1, vlcex = 0.5, title = "Alex Caruso")
radarchart(radar_data2[c(1, 2, 11), ], pcol = "darkorange", pfcol = adjustcolor("darkorange", alpha.f = 0.5), plwd = 2, cglcol = "gray", cglty = 1, vlcex = 0.5, title = "Cason Wallace")
radarchart(radar_data2[c(1, 2, 12), ], pcol = "navy", pfcol = adjustcolor("navy", alpha.f = 0.5), plwd = 2, cglcol = "gray", cglty = 1, vlcex = 0.5, title = "Jalen Suggs")
```

### Deflections v Loose Balls Recovered

```{r}
plot(x = c(hustle$Deflections[indices], 3.7), 
     y = c(hustle$Loose.Balls.Recovered[indices], 1.0), pch = 16, 
     col = c("red", "red", rep("black", times = 8), "blue"), 
     ylim = c(0.4, 1.3), xlim = c(1.5, 6.6),
     xlab = "Deflections (per game)", 
     ylab = "Loose Balls Recovered (per game)")
text(x = c(hustle$Deflections[indices], 3.7), 
     y = c(hustle$Loose.Balls.Recovered[indices], 1.0) + 0.03, 
     labels = c("Camara", "Daniels", "George", "Wembanyama", "Green", "Sengun",
                "Hart", "Caruso", "Wallace/Caruso-24", "Suggs", ""), 
     cex = 0.6)
```

### Charges Drawn v Box Outs

```{r}
plot(x = c(hustle$Charges.Drawn[indices], 0.17), 
     y = c(hustle$Box.Outs[indices], 0.4), pch = 16, 
     col = c("red", "red", rep("black", times = 8), "blue"), 
     ylim = c(0.1, 2.7),
     xlab = "Charges Drawn (per game)", 
     ylab = "Box Outs (per game)")
text(x = c(hustle$Charges.Drawn[indices] + c(0, 0.022, 0.021, 0.04, 0, 0, 0.014, 0.022, 0.024, 0.018), 0.17 + 0.03), 
     y = c(hustle$Box.Outs[indices] + c(0.12, 0, 0, 0, 0.12, -0.12, 0, 0, 0.01, -0.02), 0.4), 
     labels = c("Camara", "Daniels", "George", "Wembanyama", "Green", "Sengun",
                "Hart", "Caruso", "Wallace", "Suggs", "Caruso-24"), 
     cex = 0.6)
```

### Contested Shots v Screen Assists

```{r}
plot(x = c(hustle$Contested.Shots[indices], 5.0), 
     y = c(hustle$Screen.Assists[indices], 0.5), pch = 16, 
     col = c("red", "red", rep("black", times = 8), "blue"), ylim = c(0, 3.7),
     xlab = "Contested Shots (per game)", 
     ylab = "Screen Asists (per game)")
text(x = c(hustle$Contested.Shots[indices], 5.0), 
     y = c(hustle$Screen.Assists[indices], 0.5) + 0.13, 
     labels = c("Camara", "Daniels", "George", "Wembanyama", "Green", "Sengun",
                "Hart", "Caruso", "Wallace", "Suggs", "Caruso-24"), 
     cex = 0.6)
```

```{r}
tophustlers_table <- gt(hustle_by_rating[1:10, c(1, 2, 6:11)]) %>%
  tab_header(title = md("**Top Hustlers**"), subtitle = "Per Game Statistics")
tophustlers_table <- tophustlers_table %>%
  tab_style(style = cell_text(size = px(8)), locations = cells_body())
tophustlers_table %>%
  tab_style(style = cell_text(size = px(8)), locations = cells_column_labels())
```

## Looking at Teams

### Team Hustle Metric

```{r}
teamhustle <- read.csv("hustletotals2024-25.csv")

team_def <- teamhustle$Deflections
team_lb <- teamhustle$Loose.Balls.Recovered
team_sa <- teamhustle$Screen.Assists
team_cd <- teamhustle$Charges.Drawn
team_cs <- teamhustle$Contested.Shots
team_bo <- teamhustle$Box.Outs

team_metric <- hustle_metric(teamhustle$Deflections, teamhustle$Loose.Balls.Recovered, teamhustle$Screen.Assists, teamhustle$Contested.Shots, teamhustle$Charges.Drawn, teamhustle$Box.Outs)

barplot(team_metric, names.arg = teamhustle$TEAM, cex.names = 0.5, las = 2, 
        ylab = "Team Hustle Ratings", main = "Hustle Ratings (2024-2025)",
        col = adjustcolor("skyblue", alpha.f = 0.5), border = "skyblue")
```
### Wins v Hustle Rating

```{r}
team_wins <- c("ATL" = 26, "BKN" = 19, "BOS" = 38, "CHA" = 13, "CHI" = 22, 
               "CLE" = 44, "DAL" = 28, "DEN" = 35, "DET" = 28, "GSW" = 27, 
               "HOU" = 33, "IND" = 29, "LAC" = 29, "LAL" = 32, "MEM" = 36, 
               "MIA" = 25, "MIL" = 28, "MIN" = 30, "NOP" = 12, "NYK" = 35,
               "OKC" = 43, "ORL" = 26, "PHI" = 20, "PHX" = 26, "POR" = 23,
               "SAC" = 27, "SAS" = 23, "TOR" = 17, "UTA" = 12, "WAS" = 9)

wins_v_hustle <- lm(team_wins ~ team_metric)
summary(wins_v_hustle)

plot(x = team_metric, y = team_wins, main = "Does Hustle Impact Wins? (2024-2025)",
     ylab = "Wins", xlab = "Team Hustle Rating", pch = 16)
abline(wins_v_hustle, col = "red")
legend(x = 82, y = 17, legend = "R-Squared = 0.1413   ", cex = 0.8)
```

Based on the linear model, there seems to be little to no correlation between a team's hustle rating and the amount of games they've won.

### Wins v Deflections

```{r}
wins_v_deflections <- lm(team_wins ~ team_def)
summary(wins_v_deflections)

plot(x = team_def, y = team_wins, main = "Do Deflections Impact Wins?",
     ylab = "Wins", xlab = "Deflections", pch = 16)
abline(wins_v_deflections, col = "red")
```

### Wins v Loose Ball Recoveries

```{r}
wins_v_loose <- lm(team_wins ~ team_lb)
summary(wins_v_loose)

plot(x = team_lb, y = team_wins, 
     main = "Do Loose Ball Recoveries Impact Wins?", ylab = "Wins", 
     xlab = "Loose Balls Recovered", pch = 16)
abline(wins_v_loose, col = "red")
```

### Wins v Charges Drawn

```{r}
wins_v_charges <- lm(team_wins ~ team_cd)
summary(wins_v_charges)

plot(x = team_cd, y = team_wins, main = "Do Charges Drawn Impact Wins?", 
     ylab = "Wins", xlab = "Charges Drawn", pch = 16)
abline(wins_v_charges, col = "red")
```

### Wins v Contested Shots

```{r}
wins_v_contests <- lm(team_wins ~ team_cs)
summary(wins_v_contests)

plot(x = team_cs, y = team_wins, main = "Do Contested Shots Impact Wins?", 
     ylab = "Wins", xlab = "Contested Shots", pch = 16)
abline(wins_v_contests, col = "red")
```

### Wins v Screen Assits

```{r}
wins_v_screens <- lm(team_wins ~ team_sa)
summary(wins_v_screens)

plot(x = team_sa, y = team_wins, main = "Do Screen Assists Impact Wins?", 
     ylab = "Wins", xlab = "Screen Assists", pch = 16)
abline(wins_v_screens, col = "red")
```

### Wins v Box Outs

```{r}
wins_v_boxouts <- lm(team_wins ~ team_bo)
summary(wins_v_boxouts)

plot(x = team_bo, y = team_wins, main = "Do Box Outs Impact Wins?", 
     ylab = "Wins", xlab = "Box Outs", pch = 16)
abline(wins_v_boxouts, col = "red")
```

## Do the Top Hustlers Impact Wins?

```{r}
top2 <- function(x){
  x <- sort(x, decreasing = TRUE)
  sum(x[1:2])
}
team_top3 <- tapply(X = hustle$Hustle.Rating, INDEX = hustle$Team, top3)
team_top3
```

### Wins v Top Hustlers

```{r}
wins_v_top <- lm(team_wins ~ team_top3)
summary(wins_v_top)

plot(x = team_top3, y = team_wins, main = "Do Top Hustlers Impact Wins?", 
     ylab = "Wins", 
     xlab = "Sum of Hustle Rating of top 3 hustlers per team", 
     pch = 16)
abline(wins_v_top, col = "red")
legend(x = 82, y = 43, legend = "R-Squared = 0.1764    ", cex = 0.8)
```


```{r}
playoffs_hustle <- read.csv("teamhustleplayoffs.csv")

playoffs_hustle$Hustle.Rating <- hustle_metric(playoffs_hustle$Deflections, playoffs_hustle$Loose.Balls.Recovered, playoffs_hustle$Screen.Assists, playoffs_hustle$Contested.Shots, playoffs_hustle$Charges.Drawn, playoffs_hustle$Box.Outs)

barplot(playoffs_hustle$Hustle.Rating, names.arg = playoffs_hustle$TEAM, las = 2, 
        col = c("darkgreen", "skyblue", "lightgreen", "skyblue", "blue", "gray", "gray", 
                "gray", "gray", "blue", "gray", "skyblue", "skyblue", "gray", "gray",
                "gray"),
        ylab = "Playoff Hustle Rating")
legend(x = 3, y = 92, legend = c("Champion", "Runner-Up", "Conference Runner-Up", "Second Round Exit", "First Round Exit"), col = c("darkgreen", "lightgreen", "blue", "skyblue", "gray"), pch = 18, cex = 0.5, pt.cex = 0.8)
```

